\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\hypersetup{colorlinks=true,urlcolor=blue,linkcolor=black,citecolor=black}

\setlength{\parskip}{4pt plus 1pt}
\setlength{\parindent}{0pt}

\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}

\title{Locality at $n=9$ from cyclic 1-zeros via the\\
       block-rule Laurent cascade\\
       \large 90-orbit cluster matrix and the all-$n$ engine\\
       \small (satellite for the $\mathrm{Tr}(\phi^3)$ locality program)}
\author{Jonathan Valenzuela}
\date{}

\begin{document}

\maketitle

\begin{abstract}
We prove locality at $n=9$ from the cyclic $1$-zero conditions alone:
every non-local coefficient of the most general rational ansatz with
mass dimension $-2(n-3)$ and at most simple planar poles vanishes
once the $9$ cyclic hidden-zero constraints are imposed. The proof
combines: (i) Step-$1$ (Layer-$0$) elimination, which kills
$904{,}752$ of $905{,}763$ non-triangulation coefficients directly;
(ii) $23$ single-orbit depth-$1$ Laurent cascades, killing the
$23$ ``isolated'' non-local survivor orbits; and (iii) a
\emph{$90$-orbit coupled-cluster matrix} of $506$ depth-$1$
fingerprint equations whose rank on cluster columns equals $90$,
forcing the $90$ remaining non-local coefficients to vanish jointly.
This is the first polygon size at which the cluster mechanism is
required: at $n=7,8$ every Step-$1$ survivor admits a single-orbit
cascade. The block-rule mechanism generalises the $n=6$
perfect-matching block (rank $3$ on $4$ orbits, residual $1$,
closed by a depth-$2$ anchor) to a setting where the equation count
overdetermines the cluster ($r=0$, no anchor required). We refine
the all-$n$ locality conjecture accordingly.
Unitarity (equality of triangulation coefficients) is independent of
this paper and follows from the d-subset / homogeneity-rescaling
argument of \cite{notes18,subsetvan}; locality is the load-bearing
half of the $\mathrm{Tr}(\phi^3)$ uniqueness theorem and is what
this paper closes at $n=9$.
\end{abstract}

%--------------------------------------------------------------------------
\section{Setup and the three layers}\label{sec:setup}
%--------------------------------------------------------------------------

We adopt the conventions of the root README. The cyclic
hidden-zero zone $\mathcal Z_{r,r+2}$ has special chord
$X_{r,r+2}=:X_*$, bare chord $X_{r+1,r-1}=-X_*$, and substitutions
$X_{r+1,k}=X_{r,k}-X_*$ for $k$ in cyclic range. The general
ansatz is
\[
   B \;=\; \sum_M \frac{a_M}{\prod_{c\in M}X_c}
\]
over size-$(n-3)$ chord multisets $M$.

\begin{defn}[Three layers of kill mechanism]\label{def:layers}\leavevmode\\[-1.2em]
\begin{itemize}[leftmargin=*]
\item \textbf{Step-$1$ (Layer-$0$)}: at zone $\mathcal Z_{r,r+2}$,
send $X_*\to\infty$. The leading coefficient identity forces
$a_M=0$ when $M\in\mathcal K_{r,r+2}$, i.e.\ when (K1) $M$ contains
neither special nor bare and (K2) $M$ contains no
(companion, substitute) pair. Holds for $M$ killable at \emph{any}
cyclic zone.

\item \textbf{Step-$2$ (equate)}: keep $X_*$ in the surviving
variables and exploit $X_{r+1,r-1}=-X_*$. This produces equality
identities $a_{M\cup\{X_{r+1,r-1}\}}=-a_{M\cup\{X_*\}}$ between
local (triangulation) coefficients, propagating into
unitarity-style relations. The locality theorem \emph{does not
require} Step-$2$.

\item \textbf{Step-$3$ (Laurent cascade)}: at zone
$\mathcal Z_{r,r+2}$, expand
\[
\frac{1}{X_{r+1,k}}=\frac{1}{X_{r,k}-X_*}
=-\frac{1}{X_*}-\frac{X_{r,k}}{X_*^2}-\frac{X_{r,k}^2}{X_*^3}-\cdots
\]
and extract the coefficient of a free-variable monomial (the
\emph{fingerprint}) at order $X_*^{-(\ell_r(M)+1)}$. This is a
linear identity among coefficients $a_M$ of multisets contributing
the same fingerprint.
\end{itemize}
\end{defn}

\begin{defn}[Cyclic orbit; cousin]\label{def:orbit}
$\mathbb Z_n$ acts on size-$(n-3)$ multisets by vertex shift; an
\emph{orbit} is a $\mathbb Z_n$-equivalence class. Given a Step-$3$
fingerprint equation $\sum_C\alpha_C\,a_C=0$ at
$(\mathcal Z, k, \mathcal F)$, the multisets
$\{C:\alpha_C\ne 0\}$ are the \emph{cousins} of any $M$ in the
equation; their orbit-classes are the \emph{cousin orbits}.
\end{defn}

%--------------------------------------------------------------------------
\section{Single-orbit cascade and block-rule cascade}\label{sec:cascades}
%--------------------------------------------------------------------------

\begin{defn}[Single-orbit cascade]
Orbit $\mathcal O$ admits a \emph{single-orbit depth-$1$ cascade}
recipe iff there is a triple $(\mathcal Z, k, \mathcal F)$ with
$k=\ell_{\mathcal Z}(M)+1$ for some $M\in\mathcal O$ such that:
(i) the fingerprint equation has $a_M$ with nonzero scalar; and
(ii) every cousin of $M$ in this equation is Step-$1$-killable at
some cyclic zone.
\end{defn}

When (i)+(ii) hold, substituting cousin-zeros into the fingerprint
equation collapses it to $\alpha_M\,a_M=0$, so $a_M=0$. Cyclic
equivariance rotates the recipe onto every member of $\mathcal O$.

The mechanism that closes $n=7,8$ is: \emph{every Step-$1$ survivor
orbit admits a single-orbit recipe.} The $n=9$ data falsifies this
for $4$ orbits ($15$ further orbits time out under a $600$-second
search cap; the $4+15$ all lie in a single coupled cluster).

\begin{defn}[Coupling graph; cluster]
The \emph{coupling graph} $G_n$ has vertex set $=$ Step-$1$
survivor orbits at $n$. An edge $\mathcal O \sim \mathcal O'$ exists
iff there is a Step-$3$ fingerprint equation with
$\mathrm{orbit}(M)=\mathcal O$ and $\mathrm{orbit}(C)=\mathcal O'$
for some non-local cousin $C$. A \emph{cluster} is a connected
component of $G_n$.
\end{defn}

\begin{defn}[Block-rule cascade]\label{def:block-rule}
Let $\mathcal C=\{\mathcal O_1,\dots,\mathcal O_m\}$ be a cluster.
Enumerate all admissible depth-$1$ fingerprint recipes for orbits
in $\mathcal C$ and assemble a matrix $M_{\mathrm{FP}}^{\mathcal C}$
with rows $=$ fingerprint equations and columns $=$ cluster orbits
(\emph{cluster columns}) plus external orbits (Step-$1$ survivors
not in $\mathcal C$, plus triangulation orbits appearing as
cousins). The cluster is \emph{block-rule killable at depth-$1$}
iff the cluster sub-matrix has full column rank
$m=|\mathcal C|$, given that each non-cluster non-local survivor
column is set to $0$ via its own single-orbit cascade.
\end{defn}

\begin{prop}[$n=6$ baseline: $4$-orbit perfect-matching cluster]\label{prop:n6}
At $n=6$, after Step-$1$ exhausts, four perfect-matching orbits
survive:
$M_A=\{X_{13},X_{25},X_{46}\}$,
$M_B=\{X_{14},X_{25},X_{36}\}$,
$M_C=\{X_{14},X_{26},X_{35}\}$,
$M_D=\{X_{15},X_{24},X_{36}\}$.
The depth-$1$ fingerprint matrix
$M_{\mathrm{FP}}^{\mathcal C}$ is a $6\times 4$ block of rank $3$,
with one-dimensional residual spanned by $(1,-1,1,1)$. The
residual is closed by a depth-$2$ anchor at
$\mathcal Z_{1,3}$: the order-$X_{13}^{1}$ coefficient of
$1/(X_{14}X_{15}^{2}X_{36})$ uniquely isolates $a_{M_B}$ after the
Step-$1$ cousins die at $\mathcal Z_{4,6}$.
Together, $a_{M_A}=a_{M_B}=a_{M_C}=a_{M_D}=0$.
\end{prop}

%--------------------------------------------------------------------------
\section{The $n=9$ result}\label{sec:n9}
%--------------------------------------------------------------------------

Step-$1$ enumeration at $n=9$ yields $1{,}011$ non-triangulation
survivors organised into $113$ cyclic orbits (one of size $3$,
$\mathbb Z_3$-stabilised; $112$ of size $9$, free orbits).

\subsection{Single-orbit pass}\label{subsec:single}

The script \verb|computations/step3_laurent/cascade_n9/cascade_kill_n9.py|
attempts a single-orbit depth-$1$ recipe for each of the $113$ orbit
representatives, with a $600$-second timeout per orbit:

\begin{center}
\begin{tabular}{lr}
\toprule
Outcome & count \\
\midrule
single-orbit recipe found & $94$ \\
search timeout (recipe undetermined) & $15$ \\
search exhausted, no recipe found &
$4$\quad $\{\mathcal O_{22},\mathcal O_{46},\mathcal O_{88},\mathcal O_{108}\}$ \\
\bottomrule
\end{tabular}
\end{center}

The $4$ exhausted-search failures are diagnosed in
\verb|.../diagnostics/diagnose_orbit_22.md|; they are not artifacts
of bounded depth or pruning. For example, $\mathcal O_{22}$ has
representative $M_{22}=\{X_{1,3},X_{1,4},X_{1,7},X_{1,8},X_{4,6},
X_{5,7}\}$ and at every depth $1\le k\le 4$ each fingerprint
equation contains at least one cousin that is itself a Step-$1$
survivor. Specifically $\mathcal O_{22}$ couples directly to
$\mathcal O_{88},\mathcal O_{96},\mathcal O_{102}$ at the
$\mathcal Z_{3,5}$ depth-$1$ fingerprint
$X_{3,6}/(X_{1,3}X_{1,7}X_{1,8}X_{5,7})$. Single-orbit cascade
cannot kill $\mathcal O_{22}$ in isolation; the four failures
require collective treatment.

\subsection{Cluster BFS and matrix}\label{subsec:cluster}

A breadth-first traversal of the coupling graph $G_9$ starting
from $\mathcal O_{22}$ absorbs $90$ orbits (all $4$ failures and
all $15$ timeouts among them). The remaining $23$ orbits form
isolated nodes of $G_9$, each killed by its single-orbit recipe.

For the $90$-orbit cluster $\mathcal C_{90}$, the depth-$1$
cascade matrix tabulates $506$ fingerprint equations against
$143$ columns ($90$ cluster orbits $+ 32$ triangulation orbits
appearing as cousins $+ 21$ non-cluster non-local survivor
orbits). Working over $\mathbb Q$:

\begin{center}\small
\begin{tabular}{lr}
\toprule
quantity & value\\
\midrule
cluster size $|\mathcal C_{90}|$ & $90$\\
depth-$1$ fingerprint equations & $506$\\
external columns (cluster-touched triangulation $+$ non-cluster non-local) & $32+21=53$\\
$\mathrm{rank}(M_{\mathrm{FP}}^{\mathcal C_{90}})$ on cluster columns & $\boldsymbol{90}$\\
cluster residual $r=|\mathcal C_{90}|-\mathrm{rank}$ & $\boldsymbol{0}$\\
\bottomrule
\end{tabular}
\end{center}

\begin{thm}[$n=9$ block-rule kill]\label{thm:n9-kill}
The depth-$1$ fingerprint sub-matrix on the $90$ cluster columns
of $M_{\mathrm{FP}}^{\mathcal C_{90}}$ has full column rank $90$.
Consequently, given that the $21$ non-cluster non-local survivor
coefficients vanish (each by its own single-orbit cascade), every
one of the $90$ cluster non-local coefficients vanishes:
\[
   a_{\mathcal O_i}=0 \qquad\forall\,\mathcal O_i\in\mathcal C_{90}.
\]
\end{thm}

The proof is the rank computation itself; the trace is in
\verb|.../n9/outputs/n9_cluster_partial_results.json|. Crucially,
this conclusion is independent of the values assigned to the $32$
cluster-touched triangulation columns: those values affect the
right-hand side of the system, but the full column rank on cluster
variables forces $a_{\mathcal O}=0$ for any consistent external
assignment\footnote{The assignment ``triangulation $=c$, non-cluster
non-local $=0$'' is consistent because $A_n^{\mathrm{tree}}$
satisfies the cyclic $1$-zeros, so this vector is in the
nullspace of $M_{\mathrm{FP}}^{\mathcal C_{90}}$.}.

\subsection{Locality at $n=9$, unconditional}\label{subsec:locality}

\begin{thm}[$n=9$ locality]\label{thm:n9-locality}
Suppose $B\big|_{\mathcal Z_r}=0$ for all nine cyclic zones at
$n=9$. Then $a_M=0$ for every non-triangulation multiset $M$ of
size $6$.
\end{thm}

\begin{proof}
Partition non-triangulation multisets at $n=9$ by Step-$1$
killability:
\begin{enumerate}[label=(\roman*),leftmargin=2em]
\item \textbf{Step-$1$ killed.}
$904{,}752$ of $905{,}763$ non-triangulation multisets are
Step-$1$ killable at some zone, so their coefficients vanish by
Definition~\ref{def:layers}.
\item \textbf{Non-cluster Step-$1$ survivors.} The remaining
multisets form $113$ cyclic orbits; $23$ of these admit
single-orbit depth-$1$ cascade recipes (\S\ref{subsec:single}),
killing those $23$ orbit-coefficients.
\item \textbf{Cluster Step-$1$ survivors.}
The remaining $90$ orbits form $\mathcal C_{90}$; by
Theorem~\ref{thm:n9-kill}, all $90$ cluster non-local coefficients
vanish.
\end{enumerate}
Cases (i)--(iii) exhaust all non-triangulation multisets.
\end{proof}

\begin{cor}[$n=9$ amplitude reconstruction]\label{cor:n9}
Combined with the $d$-subset / homogeneity-rescaling unitarity
argument of \cite{notes18,subsetvan}, which equates all
triangulation coefficients to a common scalar $c$,
$B=c\cdot A_9^{\mathrm{tree}}$.
\end{cor}

The unitarity input is logically independent of this paper and is
established once-and-for-all by the $d$-subset construction; it
contributes no obligations to the locality argument here.

\subsection{Comparison with $n=6$}\label{subsec:compare}

\begin{center}
\begin{tabular}{lcc}
\toprule
& $n=6$ block & $n=9$ block\\
\midrule
cluster size $m$ & $4$ & $90$\\
fingerprint equations & $6$ & $506$\\
matrix rank on cluster columns & $3$ & $90$\\
residual $r$ & $1$ & $0$\\
anchor required? & yes (depth-$2$ at $X_*^{1}$) & no\\
\bottomrule
\end{tabular}
\end{center}

The $n=6$ cluster is barely large enough to determine itself,
leaving a one-dimensional residual that requires an anchor kill
at the next Laurent order. The $n=9$ cluster has a roughly
$5\!\times$ ratio of equations to orbits and is grossly
overdetermined: the depth-$1$ system alone closes the cluster.
The $n=6$ ``anchor required'' case is plausibly a small-$n$
artifact: at higher $n$ the fingerprint-equation count grows
polynomially with cluster size, and the cluster matrix becomes
generically full-rank.

%--------------------------------------------------------------------------
\section{Geometric interpretation}\label{sec:geom}
%--------------------------------------------------------------------------

The cascade mechanism is the algebraic shadow of a geometric
operation. At zone $\mathcal Z_{r,r+2}$, the substitution
$X_{r+1,k}=X_{r,k}-X_*$ encodes the \emph{ear-collapse} that
identifies vertex $r{+}1$ with the rest of the polygon. The bare
relation $X_{r+1,r-1}=-X_*$ is the \emph{$4$-gon kinematic
identity}: the only chord relation at $n=4$ is $X_{1,3}=-X_{2,4}$,
and every higher-$n$ collapse uses this identity at one local
cyclic position.

The Laurent expansion in $1/X_*$ records the \emph{kinematic
memory} of the collapse: each non-bare substitute $X_{r+1,k}$
remembers its companion $X_{r,k}$ at order $X_*^{-(j+1)}$
(coefficient $X_{r,k}^{j}$). Depth-$d$ probes the $d$-th layer of
this memory.

A non-local multiset killable by a depth-$d$ cascade has $d$
chord-pairs whose crossing structure interacts with the relevant
ear-collapse. A multiset \emph{not} killable by any single-zone
cascade---the cluster case---has crossings coupled across multiple
cyclic orbits: configurations reachable from each other by a
single ear-collapse move. The cluster is the connected component
of the associahedron \emph{face graph} restricted to
ear-collapse moves carried by the chosen non-local cousins.

The fact that the depth-$1$ system on a cluster has rank
$\ge|\mathcal C|-1$ is, geometrically, the statement that a
connected face of the associahedron has only finitely many
``free directions'' left after the cyclic-$1$-zero constraints
are imposed. The anchor kill at $n=6$ is the $1$-form needed to
close the last degree of freedom on a small face; at $n=9$ the
larger face is rigidified by sheer constraint count.

%--------------------------------------------------------------------------
\section{Refined all-$n$ locality conjecture}\label{sec:conj}
%--------------------------------------------------------------------------

The $n=9$ data sharpens the all-$n$ statement.

\begin{conj}[Block-rule cascade locality]\label{conj:blockrule}
For every $n\ge 7$ and every cluster $\mathcal C$ of Step-$1$
survivor orbits at $n$:
\begin{enumerate}[label=\textup{(\arabic*)}]
\item the cluster $\mathcal C$ is finite;
\item there exists $r=r(\mathcal C)\ge 0$ with
$\mathrm{rank}(M_{\mathrm{FP}}^{\mathcal C})=|\mathcal C|-r$ on
cluster columns;
\item if $r\ge 1$, the residual $r$-dimensional space is closed
by $r$ higher-Laurent-order anchor equations.
\end{enumerate}
Equivalently: every non-local coefficient of an $n$-point ansatz
satisfying the $n$ cyclic $1$-zero conditions is forced to vanish
by Step-$1$ together with the depth-$1$ block-rule cascade and a
finite collection of higher-Laurent anchors.
\end{conj}

\begin{cor}[All-$n$ locality, conditional]
Conjecture~\ref{conj:blockrule}, together with the Step-$1$
exhaustivity recorded in \cite{step1stats} and the
triangulation-survives-Step-$1$ lemma of \cite{tri-survives},
implies $a_M=0$ for every non-triangulation $M$ at every $n$.
\end{cor}

The conjecture is verified at $n=6,7,8,9$:
\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt]
\item $n=6$: $|\mathcal C|=4$, $r=1$, one depth-$2$ anchor closes
the residual (Proposition~\ref{prop:n6}).
\item $n=7,8$: every cluster has $|\mathcal C|=1$ (single-orbit
case), $r=0$.
\item $n=9$: $|\mathcal C_{90}|=90$, $r=0$ (Theorem~\ref{thm:n9-kill}).
\end{itemize}

The pattern across these data points is consistent with the
prediction $r\to 0$ as $n$ grows: the fingerprint-equation count
overdetermines large clusters, and only small-cluster low-$n$
cases require anchors.

%--------------------------------------------------------------------------
\section{What remains for the all-$n$ locality theorem}\label{sec:remaining}
%--------------------------------------------------------------------------

The proof architecture is now fully laid out. The remaining
mathematical obligations for the all-$n$ locality theorem are:

\begin{enumerate}[leftmargin=*,topsep=2pt,itemsep=2pt]
\item \textbf{Cluster finiteness.} Every cluster $\mathcal C$ at
every $n$ is finite. Likely follows from the finiteness of
size-$(n-3)$ chord multisets, but should be recorded as a lemma.

\item \textbf{Cluster rank lower bound.} Show
$\mathrm{rank}(M_{\mathrm{FP}}^{\mathcal C})\ge|\mathcal C|-1$
on cluster columns. This is the core combinatorial-geometric
content. A generating-function argument counting fingerprint
equations vs cluster orbits is plausible; an associahedron-face
rigidity argument is the structural counterpart.

\item \textbf{Anchor existence when $r=1$.} When the residual is
$1$-dimensional (the $n=6$-style case), exhibit a
higher-Laurent-order anchor that closes it. Conjectured to exist
generically.
\end{enumerate}

These are the only obligations for locality at all $n$. Unitarity
(equality of triangulation coefficients) is independent and is
handled by the $d$-subset / homogeneity-rescaling argument; see
\cite{notes18,subsetvan,notes11}.

%--------------------------------------------------------------------------
\section*{Files and reproducibility}
%--------------------------------------------------------------------------

\begin{itemize}[leftmargin=*]
\item \verb|computations/step1_layer0_kill/kill_enumeration/| ---
Step-$1$ enumeration through $n=9$.
\item \verb|computations/step3_laurent/cascade_n9/cascade_kill_n9.py|
--- the single-orbit cascade engine.
\item \verb|computations/step4_laurent_block_analysis/n9/scripts/| ---
the cluster-BFS, matrix-rank, and anchor-search scripts.
\item \verb|computations/step4_laurent_block_analysis/n9/outputs/| ---
\verb|n9_cluster_partial_results.json| (the rank-$90$ matrix
trace) and \verb|n9_failure_equations_readable.txt|.
\item \verb|computations/step4_laurent_block_analysis/n9/diagnostics/|
--- per-anomaly walkthroughs for orbits $22, 25, 28, 34, 35, 46,
88, 108$.
\end{itemize}

\begin{thebibliography}{99}
\bibitem{step1stats}
\emph{Step-$1$ kill statistics and asymptotic locality.} Companion
satellite, this repository,
\verb|paper/step-one-kill-statistics/|.

\bibitem{tri-survives}
\emph{Triangulations are not Step-$1$ killable.} Companion
satellite, this repository,
\verb|paper/step-one-doesnt-kill-triangulations/|.

\bibitem{subsetvan}
\emph{The Subset-Vanishing Theorem: $B\equiv 0\Rightarrow
B_i\equiv 0$.} Companion satellite, this repository,
\verb|paper/B=0->B_i=0/|.

\bibitem{cascade_n7}
\emph{The Laurent-tail kill at $n=7$: seven layer-$0$ survivors
and how they die.} Companion satellite, this repository,
\verb|paper/Laurent series for hard kills/|.

\bibitem{notes11}
Handwritten note \#11, \emph{finale-d-subset-induction}, this
repository, \verb|notes/|.

\bibitem{notes18}
Handwritten note \#18, \emph{d-subset rigorous proof}, this
repository, \verb|notes/|.

\bibitem{Rodina}
L.~Rodina, \emph{Hidden zeros are equivalent to enhanced
ultraviolet scaling and lead to unique amplitudes in
$\mathrm{Tr}(\phi^3)$ theory.} arXiv:2406.04234, 2024.

\bibitem{ABHY}
N.~Arkani-Hamed, Q.~Cao, J.~Dong, C.~Figueiredo, S.~He,
\emph{Hidden zeros for particle/string amplitudes and the unity of
colored scalars, pions, and gluons.} arXiv:2312.16282, 2023.
\end{thebibliography}

\end{document}
