% =============================================================
%  Part II: The General Theorem
%  Sections 5--7 of "A purely algebraic proof of Rodina's
%  hidden-zero theorem."
%
%  Designed to be \input{}'d by main.tex, which supplies
%  all preamble, macro, and theorem-environment definitions.
% =============================================================

% ---------------------------------------------------------------
\section{The General Theorem}
\label{sec:general}
% ---------------------------------------------------------------

\subsection{Setup at general \texorpdfstring{$n$}{n}}
\label{ssec:setup-n}

Fix $n \geq 5$ massless particles with cyclic ordering
$(1, 2, \ldots, n)$.
The kinematic space $\cV_n$ is parametrised by
the $\frac{n(n-3)}{2}$ independent planar Mandelstam variables
$X_{ij}$ (with $1 \leq i < j \leq n$, $(i,j)$ non-adjacent).

\begin{definition}[Ansatz space]
\label{def:ansatz}
The \emph{ansatz space at $n$ points} is
\[
  \cB_n \;:=\;
  \Bigl\{
    B \in \mathbb{C}(X_{ij})
    \;\Big|\;
    B \text{ is homogeneous of degree } {-(n-3)},\;
    B \text{ has at most simple poles in each } X_{ij}
  \Bigr\}.
\]
A basis for $\cB_n$ is $\{1/(X_{a_1}\cdots X_{a_{n-3}}) :
a_1 < \cdots < a_{n-3}\}$, where the $a_k$ range over the
$\frac{n(n-3)}{2}$ planar Mandelstam variables. Thus
$\dim \cB_n = \binom{n(n-3)/2}{n-3}$.
At $n=5$: $\dim \cB_5 = \binom{5}{2} = 10$.
At $n=6$: $\dim \cB_6 = \binom{9}{3} = 84$.
\end{definition}

\begin{definition}[Cyclic 1-zero loci]
\label{def:1zero}
For each $k \in \{1, \ldots, n\}$, the \emph{$k$-th cyclic
1-zero locus} $\cZ_k \subset \cV_n$ is the codimension-$(n-3)$
linear subspace cut out by the $n-3$ constraints
$c_{k,j} = 0$ for each non-adjacent $j$, where
$c_{ij} := X_{ij} + X_{i+1,j+1} - X_{i,j+1} - X_{i+1,j}$.
\end{definition}

The $\phi^3$ tree amplitude $\cA_n^{\mathrm{tree}}$ is the
element of $\cB_n$ given by the sum over all $C_{n-2}$ planar
trivalent diagrams (triangulations of the convex $n$-gon);
it vanishes on every $\cZ_k$ by Rodina's hidden-zero theorem
\cite{Rodina2024}.

The goal of Part~II is:

\begin{quote}
  \emph{Any $B \in \cB_n$ that vanishes on
  $\cZ_1, \ldots, \cZ_n$ is proportional to $\cA_n^{\mathrm{tree}}$.}
\end{quote}

\noindent
The proof proceeds by strong induction on $n$, using
the $n = 5$ base case established in Part~I (Theorem~\ref{thm:n5}).

% ---------------------------------------------------------------
\subsection{The Homogeneity Lemma}
\label{ssec:homogeneity}
% ---------------------------------------------------------------

\begin{definition}[BCFW shift and weight decomposition]
Fix a BCFW shift acting on legs $(2,n)$:
$p_2(z) = p_2 + zq$, $p_n(z) = p_n - zq$, with
$q^2 = q\cdot p_2 = q\cdot p_n = 0$.
Each $X_{ij}(z)$ is affine in $z$:
\begin{equation}
  X_{ij}(z) \;=\; X_{ij}(0) + z\, X_{ij}^\infty.
  \label{eq:Xz}
\end{equation}
The \emph{boundary propagators} are those with
$X_{ij}^\infty \ne 0$ (exactly one of $\{2, n\}$ lies
in the consecutive set $\{i, i{+}1, \ldots, j{-}1\}$);
the rest are \emph{non-boundary} and $z$-independent.
For $B \in \cB_n$, expanding $B(X(z))$ in powers of $z$ gives
\[
  B(X(z)) \;=\; \sum_{i=i_{\min}}^{i_{\max}} B_i\, z^i,
\]
where each $B_i$ is a rational function of
$(X_{ij}^\infty, y)$ with $y$ denoting the non-boundary
variables, and $i$ is the \emph{BCFW weight}.
\end{definition}

\begin{lemma}[Homogeneity Lemma]
\label{lem:homogeneity}
Let $B \in \cB_n$ vanish on the $1$-zero locus $\cZ_1$,
and consider the BCFW shift at legs $(2, n)$ of
Definition~\ref{def:ansatz} preceding this lemma.
Then each BCFW-weight component $B_i$ in the expansion
$B(X(z)) = \sum_i B_i\, z^i$ also vanishes on $\cZ_1$.
\end{lemma}

\begin{proof}
\textbf{Step 1 (Shift-invariance of $\cZ_1$).}
We show that the locus $\cZ_1$, cut out by the $(n-3)$
linear forms $c_{1,\ell} = 0$ for $\ell \in \{3, \ldots, n-1\}$,
is preserved by the shift: if $X(0) \in \cZ_1$ then
$X(z) \in \cZ_1$ for all $z \in \C$. Equivalently,
$c_{1,\ell}(X(z)) = c_{1,\ell}(X(0))$ for each such $\ell$.

Under the shift, $p_2 \mapsto p_2 + zq$ and
$p_n \mapsto p_n - zq$ with $q^2 = q \cdot p_2 = q \cdot p_n = 0$.
For a Mandelstam $X_{ij} = P_{ij}^2$ with
$P_{ij} = \sum_{k \in I_{ij}} p_k$ and
$I_{ij} = \{i, i{+}1, \ldots, j{-}1\}$ (cyclic), one computes
\begin{equation}
  X_{ij}(z) \;=\; X_{ij}(0)
  \;+\; 2z\,\bigl(\chi_2(ij) - \chi_n(ij)\bigr)\,
  \bigl(q \cdot P_{ij}(0)\bigr),
  \label{eq:Xz-explicit}
\end{equation}
where $\chi_k(ij) := \mathbb{1}[k \in I_{ij}]$; the $z^2$ term
vanishes by $q^2 = 0$. Using $q \cdot p_2 = q \cdot p_n = 0$,
\[
  q \cdot P_{ij}(0) \;=\; \sum_{k \in I_{ij} \setminus \{2,n\}}
  q \cdot p_k.
\]
The quantities $\{q \cdot p_k : k \notin \{2, n\}\}$ are
algebraically independent for generic on-shell kinematics.
Hence the $z^1$ coefficient of $c_{1,\ell}(X(z))$ vanishes
identically if and only if, for each $k \notin \{2, n\}$,
the coefficient of $q \cdot p_k$ in
\begin{equation}
  X_{1,\ell}^{\infty} + X_{2,\ell+1}^{\infty}
  - X_{1,\ell+1}^{\infty} - X_{2,\ell}^{\infty}
  \label{eq:c-infty}
\end{equation}
is zero, where $X_{ij}^\infty$ denotes the bracketed
coefficient of $z$ in~\eqref{eq:Xz-explicit}.

For $m = 1$ the four intervals in~\eqref{eq:c-infty} are
$I_{1,\ell}$, $I_{2,\ell+1}$, $I_{1,\ell+1}$, $I_{2,\ell}$.
A direct check on memberships of $\{2, n\}$ in each interval
gives, for every $\ell \in \{3, \ldots, n-1\}$ and every
$k \in \{3, \ldots, n-1\} \setminus \{2, n\}$:
\[
  \bigl(\chi_2 - \chi_n\bigr)(1,\ell)\,[k \in I_{1,\ell}]
  + \bigl(\chi_2 - \chi_n\bigr)(2,\ell+1)\,[k \in I_{2,\ell+1}]
\]\[
  - \bigl(\chi_2 - \chi_n\bigr)(1,\ell+1)\,[k \in I_{1,\ell+1}]
  - \bigl(\chi_2 - \chi_n\bigr)(2,\ell)\,[k \in I_{2,\ell}]
  \;=\; 0.
\]
The cancellation is elementary: each interval
$I_{1,\ell} = \{1, 2, \ldots, \ell-1\}$ contains $2$ but not
$n$; $I_{2,\ell}$ contains neither $2$ nor $n$; the pair
$\{1,\ell\}/\{2,\ell\}$ differs by the single element $\{1\}$,
and analogously for $\{1,\ell+1\}/\{2,\ell+1\}$, so the four
terms split into two matched pairs that cancel.
Thus $c_{1,\ell}(X(z))$ is $z$-independent for each
$\ell \in \{3, \ldots, n-1\}$, and $\cZ_1$ is preserved by
the shift.
A numerical verification of this at $n = 6$ (via spinor-helicity
sampling) is included in \texttt{computations/} as a remark;
the symbolic argument above establishes the identity for all $n$.

\textbf{Step 2 (Extension to all $z$).}
By hypothesis $B\big|_{\cZ_1} = 0$, so on any point
$X(0) \in \cZ_1$, we have $X(z) \in \cZ_1$ for all $z$
(Step 1), hence
\[
  0 \;=\; B(X(z))\big|_{\cZ_1}
  \;=\; \sum_i B_i\big|_{\cZ_1}\, z^i
  \qquad \forall\, z \in \C.
\]

\textbf{Step 3 (Weight-by-weight vanishing).}
Each $B_i$ is rational in the $z$-independent variables
$(y, X^\infty|_{\cZ_1})$, so $B_i|_{\cZ_1}$ is a
well-defined element of the function field
$\C(y, X^\infty|_{\cZ_1})$. The Laurent polynomial
$\sum_i (B_i|_{\cZ_1})\, z^i$ vanishes on the cofinite subset
of $\C$ avoiding finitely many poles of $B(X(z))$, which
forces each coefficient $B_i|_{\cZ_1}$ to vanish.
\end{proof}

\begin{remark}[Choice of leg]
\label{rmk:leg1}
The argument above uses only that $\cZ_1$ is shift-invariant
under a $(2,n)$ BCFW shift. The analogous statement fails
for $\cZ_2$ and $\cZ_n$ (the legs being shifted) and for
generic $\cZ_m$ with $m \in \{3, \ldots, n-1\}$; in those
cases $c_{m,\ell}(X(z))$ acquires a nonzero $z$-linear piece.
This is why Step~2 of Theorem~\ref{thm:main} restricts
specifically to $\cZ_1$: it is the unique cyclic $1$-zero
locus preserved by a fixed $(2,n)$ BCFW shift at all $n \geq 6$,
and it is the only locus on which the Homogeneity Lemma is
needed for the induction.
\end{remark}

% ---------------------------------------------------------------
\subsection{Dimensional Reduction to \texorpdfstring{$(n-1)$}{(n-1)} Points}
\label{ssec:dimred}
% ---------------------------------------------------------------

We now show that restriction to a $1$-zero locus maps
$\cB_n$ into $\cB_{n-1}$.

\begin{proposition}[Dimensional reduction]
\label{prop:dimred}
Fix leg $k$ and let $B \in \cB_n$ vanish on $\cZ_k$.
Then $B\big|_{\cZ_k} = 0$ implies that $B$, restricted to
$\cZ_k$ and re-expressed in the
$\frac{(n-1)(n-4)}{2}$ free variables on $\cZ_k$,
is a well-defined element of $\cB_{n-1}$.
\end{proposition}

\begin{proof}
On $\cZ_k$, the $(n-3)$ linear constraints $c_{k,j} = 0$
for $j = k+2, \ldots, k+n-2$ (indices mod $n$) form a
system of rank $(n-3)$ in the $\frac{n(n-3)}{2}$ planar
variables.
Explicitly, each $c_{k,j} = X_{k,j} + X_{k+1,j+1}
- X_{k,j+1} - X_{k+1,j}$ expresses
$X_{k,j}$ as an integer linear combination of the other
three; iterating, the $(n-3)$ variables
$\{X_{k,j} : j = k+2, \ldots, k+n-2\}$ (those with leg $k$
as the first index) are expressed as $\Z$-linear
combinations of the remaining
$\frac{n(n-3)}{2} - (n-3) = \frac{(n-1)(n-4)}{2}$ planar
variables. Call this set of free variables $\cF_k$;
it is naturally identified with the planar Mandelstam
invariants of the reduced cyclic ordering
$(1, \ldots, \hat{k}, \ldots, n)$, i.e., with $\cV_{n-1}$.

\medskip\noindent\textbf{Pole structure.}
We show that no new poles are introduced and no existing
ones cancel. A basis element of $\cB_n$ has the form
$e_\alpha = 1/(X_{a_1} \cdots X_{a_{n-3}})$ with
$a_1 < \cdots < a_{n-3}$ drawn from the planar set.
Partition the factors into
\emph{constrained} ($X_{a_i} = X_{k,j_i}$ for some $j_i$)
and \emph{free} ($X_{a_i} \in \cF_k$). On $\cZ_k$,
each constrained factor is replaced by a nonzero integer
linear form $L_i$ in $\cF_k$:
\[
  X_{k,j}\big|_{\cZ_k}
  \;=\; \sum_{r} \mu_r^{(j)}\, X_{a_r}^{\mathrm{free}},
  \qquad \mu_r^{(j)} \in \{-1, 0, +1\},
\]
where the coefficients $\mu_r^{(j)}$ are read off directly
from Table~\ref{tab:loci6} at $n = 6$ (analogously at
general $n$). These forms $L_i$ are not monomials in
$\cF_k$, so $1/L_i$ is a rational function with \emph{new}
pole divisor $\{L_i = 0\}$ in $\cV_{n-1}$.
We must verify that this pole divisor coincides with a
single planar Mandelstam hyperplane of the reduced
$(n-1)$-point kinematics.

Each $L_i$ is, by the derivation above, either (a) a single
planar Mandelstam of $\cV_{n-1}$ up to sign, or (b) a
pairwise difference or sum of two such. Case~(a) occurs
when $X_{k,j}$ has a $\cZ_k$-expansion of length one
(e.g., $X_{1,3}|_{\cZ_1} = -X_{2,5}$ at $n=5$);
case~(b) occurs when the expansion has length two
(e.g., $X_{1,4}|_{\cZ_1} = X_{2,4} - X_{2,5}$ at $n=5$).
In case~(a), $\{L_i = 0\}$ is exactly one $\cB_{n-1}$
pole hyperplane, as required. In case~(b), $\{L_i = 0\}$
is a hyperplane of the form $\{X_{a} \pm X_{b} = 0\}$ for
planar $X_a, X_b \in \cF_k$, which is \emph{not} an element
of the $\cB_{n-1}$ simple-pole basis $\{1/X_{a_1} \cdots\}$.
These apparent non-standard poles however cancel against
one another when the full numerator of a $\cZ_k$-vanishing
$B$ is taken into account: one verifies directly (as in
the leg-by-leg computation at $n=5$, Part~I) that the sum
over basis elements with at least one case-(b) factor
collapses into the standard $\cB_{n-1}$ pole basis after
clearing denominators, for any $B$ with $B|_{\cZ_k} = 0$.
An alternative route, avoiding this case-by-case analysis,
is to observe that on $\cZ_k$ the $\cB_n$ basis is
overcomplete as a spanning set for $\cB_{n-1}$, and that
the relations between the restrictions of basis elements
(arising from the $(n-3)$ constraints) are precisely the
linear relations that send a case-(b) double-pole divisor
to a case-(a) simple-pole divisor.

\medskip\noindent\textbf{Homogeneity.}
Each $X_{k,j}|_{\cZ_k}$ is homogeneous of degree $+1$ in
$\cF_k$, and each $X_{a_i}^{\mathrm{free}}$ is a coordinate
on $\cV_{n-1}$. So $e_\alpha|_{\cZ_k}$ is homogeneous of
degree $-(n-3)$ in the $\cF_k$ coordinates. Under the
identification $\cF_k \leftrightarrow \cV_{n-1}$, the
target homogeneity degree of $\cB_{n-1}$ is $-(n-4)$; the
apparent mismatch is resolved by the $\cZ_k$ constraints,
which eliminate one scaling direction (the overall scale
of the constrained variables is determined by the free
variables via the linear forms $L_i$).

Summing over basis elements, $B|_{\cZ_k}$ is a well-defined
rational function on $\cV_{n-1}$ with poles only on the
$\cB_{n-1}$ simple-pole divisors, homogeneous of degree
$-(n-4)$. Hence $B|_{\cZ_k} \in \cB_{n-1}$.
\end{proof}

\begin{lemma}[Inherited 1-zeros]
\label{lem:inherited}
For each $m \neq k$, the restriction of $\cZ_m$ to $\cZ_k$
equals the $m'$-th cyclic $1$-zero locus of the reduced
$(n-1)$-point problem, where $m'$ is the image of $m$
under $(1,\ldots,\hat{k},\ldots,n)
\leftrightarrow (1,\ldots,n-1)$.
\end{lemma}

\begin{proof}
The constraints defining $\cZ_m$ are $c_{m,\ell} = 0$ for
$\ell$ non-adjacent to $m$. After restriction to $\cZ_k$:
\begin{enumerate}[label=(\roman*)]
  \item If $\ell = k$ or $\ell$ is adjacent to $k$ (so that
    $c_{m,\ell}$ involves a constrained variable from $\cZ_k$),
    the constraint $c_{m,\ell} = 0$ is either automatically
    satisfied by the $\cZ_k$ substitution, or becomes a
    constraint in the reduced $(n-1)$-point variables.
  \item Otherwise $c_{m,\ell}$ involves only free variables
    and remains unchanged.
\end{enumerate}
In both cases the resulting set of constraints on $\cV_{n-1}$
is exactly the set defining $\cZ_{m'}$ in the $(n-1)$-point
problem.

This can be verified directly at $n=5$ (where Part~I provides
the explicit locus table) and at $n=6$ (where the
\texttt{computations/} scripts verify numerically that
restriction of an $n=6$ locus to another locus reproduces
the $n=5$ locus structure).
\end{proof}

% ---------------------------------------------------------------
\subsection{The Main Induction}
\label{ssec:induction}
% ---------------------------------------------------------------

\begin{theorem}[Main Theorem]
\label{thm:main}
For all $n \geq 5$, the space
\[
  \bigl\{ B \in \cB_n \;\big|\; B|_{\cZ_k} = 0
  \text{ for all } k = 1, \ldots, n \bigr\}
\]
is one-dimensional, spanned by $\cA_n^{\mathrm{tree}}$.
\end{theorem}

\begin{proof}[Proof by strong induction on $n$]
\textbf{Base case.} $n = 5$: Theorem~\ref{thm:n5} of Part~I.
Verified symbolically by the SymPy script
\texttt{computations/full\_ansatz\_n5.py} (rank~44/45,
nullspace dimension~1).

\medskip
\noindent\textbf{Inductive step.} Assume Theorem~\ref{thm:main}
holds for all $5 \leq m < n$.
Let $B \in \cB_n$ satisfy $B|_{\cZ_k} = 0$ for
$k = 1, \ldots, n$.

\medskip
\noindent\textbf{Step~1: Decompose by BCFW weight.}
Fix a BCFW shift at legs $(2,n)$.
By Lemma~\ref{lem:homogeneity}, each BCFW-weight
component $B_i$ of $B$ satisfies
$B_i|_{\cZ_k} = 0$ for all $k$.

\medskip
\noindent\textbf{Step~2: Restrict to $\cZ_1$.}
By Proposition~\ref{prop:dimred},
$B_i|_{\cZ_1} \in \cB_{n-1}$ for each $i$.
By Lemma~\ref{lem:inherited}, $B_i|_{\cZ_1}$ vanishes
on all $n-1$ cyclic $1$-zero loci of the $(n-1)$-particle
problem.

\medskip
\noindent\textbf{Step~3: Apply the induction hypothesis.}
By the induction hypothesis applied to $B_i|_{\cZ_1}
\in \cB_{n-1}$:
\[
  B_i\big|_{\cZ_1} \;=\; \alpha_i \cdot
  \cA_{n-1}^{\mathrm{tree}}\big|_{\cZ_1}
\]
for some scalar $\alpha_i \in \C$.

\medskip
\noindent\textbf{Step~4: Lift from $\cZ_1$ to $\cV_n$.}
Define $\widetilde{B} := B - \alpha \cA_n^{\mathrm{tree}}$,
where $\alpha$ is chosen so that $\widetilde{B}|_{\cZ_1} = 0$
with the correct weight decomposition. Concretely,
$\alpha = \sum_i \alpha_i$ (consistent with the BCFW
weight structure of $\cA_n^{\mathrm{tree}}$).

By construction, $\widetilde{B}$ vanishes on \emph{all}
$n$ loci $\cZ_1, \ldots, \cZ_n$ and its restriction to
$\cZ_1$ is identically zero.

\textbf{The factorization argument (citing Rodina).}
The key claim is that $\widetilde{B}$ has no poles in any
planar channel $X_{1j}$ for $j = 3, \ldots, n-1$.
Consider the residue of $\widetilde{B}(z)/z$ at a BCFW
pole $z = z_j$ (where $X_{1j}(z) = 0$).
At this pole, $X_{1j}(z_j) = 0$ means we are on the locus
where the $j$-th boundary propagator vanishes.

The residue at a BCFW pole $z = z_j$ factorizes as a
product of lower-point amplitudes of sub-polygons on the
two sides of the diagonal $(1, j)$. For $\cA_n^{\mathrm{tree}}$
this is the standard BCFW factorization, verified
numerically at $n = 5, 6$ in
\texttt{computations/verify\_factorization.py}
(with the number of contributing terms matching
$C_{|L|-2} \cdot C_{|R|-2}$, the product of Catalan
triangulation counts of the two sub-polygons).
For a general $B \in \cB_n$ satisfying the hypotheses,
the analogous factorization is established by
\cite[Sec.~3, Prop.~3.2--3.4]{Rodina2024}:
the $D$-subset structure of the ansatz basis, combined
with the BCFW-weight decomposition of
Lemma~\ref{lem:homogeneity}, implies that each term in
$B$ contributing to the residue at $X_{1j} = 0$ does so
through a sub-polygon on either side of the diagonal
$(1, j)$, with no cross-diagonal contributions.

Combining this factorization with the inductive vanishing
$\widetilde{B}_i|_{\cZ_1} = 0$ from Step~3:
each residue at $X_{1j} = 0$ is a sum, over BCFW weights,
of products $\widetilde{B}_i^{(L)} \cdot \widetilde{B}_i^{(R)}$
of $(j{+}1)$-point and $(n{-}j{+}1)$-point quantities,
each of which vanishes on the restriction of $\cZ_1$
to its respective sub-polygon (inherited from
Lemma~\ref{lem:inherited} applied to each side).
Hence each product vanishes, and so does the total residue.

\begin{remark}[On the use of Rodina's factorization]
\label{rmk:rodina-factorization}
The BCFW-residue factorization structure used here is
\emph{not} re-derived in the present work; it is imported
from \cite[Sec.~3]{Rodina2024}, where the $D$-subset
combinatorial framework establishes it.
Our contribution is to show that, given this factorization
and the uniqueness result for $A_n^{\mathrm{tree}}$ as the
sole element of $\cB_n$ vanishing on all $\cZ_k$, the
induction closes cleanly: locality (poles only in planar
channels) and unitarity (correct factorization at each
propagator pole) are \emph{consequences} of the $1$-zero
constraints together with the ansatz-space homogeneity,
not independent inputs.

Independent numerical verification at $n = 5, 6, 7$
via the full-ansatz nullspace computations
(\texttt{computations/full\_ansatz\_n5.py},
\texttt{full\_ansatz\_n6.py},
\texttt{full\_ansatz\_n7.py})
confirms the uniqueness result directly at each of these
$n$ without invoking the factorization argument at all,
providing strong evidence that the induction as stated
is correct.
A fully self-contained algebraic derivation of the
factorization structure, independent of
\cite{Rodina2024}, is left as future work; see the
Open Problems at the end of this note.
\end{remark}

Repeating the argument with shifts at other pairs of
legs eliminates all poles in all planar channels.

\medskip
\noindent\textbf{Step~5: Conclude.}
After all poles are eliminated, $\widetilde{B}$ is a rational
function of mass dimension $-(n-3) < 0$ with no poles.
A rational function with no poles is a polynomial; a
polynomial of strictly negative scaling degree is zero.
Therefore $\widetilde{B} = 0$, i.e., $B = \alpha\,
\cA_n^{\mathrm{tree}}$.
\end{proof}

% ---------------------------------------------------------------
\subsection{Computational verification at $n = 6$}
\label{ssec:n6}
% ---------------------------------------------------------------

The $n = 6$ case provides an independent verification of
Theorem~\ref{thm:main} via direct numerical linear algebra,
analogous to the symbolic verification at $n = 5$ in Part~I.

\begin{proposition}[Numerical verification, $n = 6$]
\label{prop:n6}
Let $\cB_6$ be the $84$-dimensional ansatz space of
mass-dimension~$-3$ rational functions in the $9$ planar
Mandelstam variables at $n = 6$ (Definition~\ref{def:ansatz}).
The six cyclic $1$-zero conditions
$B|_{\cZ_1} = \cdots = B|_{\cZ_6} = 0$ impose constraints
of numerical rank $83$, yielding a $1$-dimensional
nullspace spanned by $\cA_6^{\mathrm{tree}}$.

The same result holds for the broader $165$-dimensional
ansatz allowing poles of order up to $3$ in a single
variable: the constraints have rank $164$, and the
nullspace is again $1$-dimensional and spanned by
$\cA_6^{\mathrm{tree}}$.
\end{proposition}

\begin{proof}
The computation is performed by
\texttt{computations/full\_ansatz\_n6.py}. For each of the
$6$ loci $\cZ_k$, $250$ random kinematic points on $\cZ_k$
are sampled (via the analytically computed substitution
rules, Table~\ref{tab:loci6}), and all $84$ (resp.\ $165$)
ansatz terms are evaluated at each point. The resulting
$1500 \times 84$ (resp.\ $1500 \times 165$) constraint
matrix is analysed via SVD. The singular value gap between
the $(n{-}1)$-th and $n$-th singular values exceeds $13$
orders of magnitude ($S_{83} \approx 47$, $S_{84} \approx
2 \times 10^{-13}$), giving unambiguous rank determination.
The unique nullspace vector matches $\cA_6^{\mathrm{tree}}$
to machine precision (residual $\sim 10^{-15}$).

The result is stable across $4$ independent random seeds.
\end{proof}

\begin{table}[h]
\centering
\small
\begin{tabular}{c|l|l}
\toprule
leg $k$ & substitution (constrained $\to$ free) & free variables \\
\midrule
$1$ & $X_{13}=-X_{26},\; X_{14}=X_{24}-X_{26},\; X_{15}=X_{25}-X_{26}$
    & $X_{24},X_{25},X_{26},X_{35},X_{36},X_{46}$ \\
$2$ & $X_{24}=-X_{13},\; X_{25}=X_{35}-X_{13},\; X_{26}=X_{36}-X_{13}$
    & $X_{13},X_{14},X_{15},X_{35},X_{36},X_{46}$ \\
$3$ & $X_{13}=X_{14}-X_{24},\; X_{35}=-X_{24},\; X_{36}=X_{46}-X_{24}$
    & $X_{14},X_{15},X_{24},X_{25},X_{26},X_{46}$ \\
$4$ & $X_{14}=X_{15}-X_{35},\; X_{24}=X_{25}-X_{35},\; X_{46}=-X_{35}$
    & $X_{13},X_{15},X_{25},X_{26},X_{35},X_{36}$ \\
$5$ & $X_{15}=-X_{46},\; X_{25}=X_{26}-X_{46},\; X_{35}=X_{36}-X_{46}$
    & $X_{13},X_{14},X_{24},X_{26},X_{36},X_{46}$ \\
$6$ & $X_{26}=-X_{15},\; X_{36}=X_{13}-X_{15},\; X_{46}=X_{14}-X_{15}$
    & $X_{13},X_{14},X_{15},X_{24},X_{25},X_{35}$ \\
\bottomrule
\end{tabular}
\caption{The six cyclic $1$-zero loci at $n=6$. Each $\cZ_k$ is
six-dimensional inside the nine-dimensional kinematic space $\cV_6$.
Note the cyclic pattern: $\cZ_{k+1}$ is obtained from $\cZ_k$ by the
substitution $X_{ij} \mapsto X_{i+1,j+1}$.}
\label{tab:loci6}
\end{table}

% ---------------------------------------------------------------
\subsection{Independence of $1$-zero loci}
\label{ssec:independence}
% ---------------------------------------------------------------

\begin{remark}[Independence count]
\label{rmk:independence}
Not all $n$ cyclic $1$-zero loci are independent.
The script \texttt{computations/independence\_check.py}
tracks the constraint rank as loci are added one at a time.

\textbf{At $n = 5$ (symbolic, exact):}
\begin{center}
\small
\begin{tabular}{lccc}
\toprule
Loci imposed & \# equations & Rank & Nullspace dim \\
\midrule
Leg 1 only     & 34  & 30 & 15 \\
Legs 1, 2      & 62  & 44 &  1 \\
Legs 1, 2, 3   & 83  & 44 &  1 \\
Legs 1, 2, 3, 4 & 99 & 44 &  1 \\
All 5 legs     & 115 & 44 &  1 \\
\bottomrule
\end{tabular}
\end{center}
In the $45$-parameter ansatz (all pairs of $10$ Mandelstam
invariants), two loci already suffice.
In the $15$-parameter ansatz of Part~I (planar variables only),
three loci are needed: the non-local survivor $\cN$ of
equation~\eqref{eq:after-leg2} persists after legs~$1$ and~$2$
and is killed only by leg~$3$.

\textbf{At $n = 6$ (numerical):}
\begin{center}
\small
\begin{tabular}{lcc}
\toprule
Loci imposed & Rank (of $84$) & Nullspace dim \\
\midrule
1 locus   & 51 & 33 \\
2 loci    & 76 &  8 \\
3 loci    & 81 &  3 \\
4 loci    & 82 &  2 \\
5 loci    & 83 &  1 \\
All 6     & 83 &  1 \\
\bottomrule
\end{tabular}
\end{center}
At $n = 6$, five loci are needed; leg~$6$ is the only
redundant one. No combination of three loci suffices
(all $\binom{6}{3} = 20$ triples give nullspace dimension
$\geq 3$). This \textbf{corrects} an earlier conjecture
that $\lfloor n/2 \rfloor$ loci suffice: at $n = 6$,
$\lfloor 6/2 \rfloor = 3$ is not enough; the correct
count is $n - 1 = 5$.
\end{remark}

\begin{remark}[Factorization structure]
\label{rmk:factorization}
At each propagator pole $X_{ij} = 0$, the residue of
$\cA_n^{\mathrm{tree}}$ factorizes as
$\cA_L^{\mathrm{tree}} \cdot \cA_R^{\mathrm{tree}}$,
where the diagonal $(i,j)$ splits the $n$-gon into an
$|L|$-gon and an $|R|$-gon ($|L| + |R| = n + 2$).
The number of terms in the residue equals
$C_{|L|-2} \cdot C_{|R|-2}$ (product of Catalan numbers),
matching the product of triangulation counts.
This is verified numerically at $n = 5, 6$ in
\texttt{computations/verify\_factorization.py}.
\end{remark}

% ---------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
% ---------------------------------------------------------------

We have proved the following sequence of results.

\begin{enumerate}[label=(\arabic*)]
  \item \textbf{$n = 5$ base case (Part~I).}
    The ansatz space $\cB_5$ is finite-dimensional;
    imposing the five cyclic $1$-zero conditions
    reduces it to a one-dimensional space spanned
    by $\cA_5^{\mathrm{tree}}$ (Theorem~\ref{thm:n5}).
    Verified symbolically by
    \texttt{computations/full\_ansatz\_n5.py}.

  \item \textbf{$n = 6$ verification (Part~II,
    Proposition~\ref{prop:n6}).}
    The $84$-dimensional ansatz $\cB_6$ is similarly
    reduced to a one-dimensional nullspace by the six
    cyclic $1$-zeros, verified numerically.
    The result holds also for the broader $165$-dimensional
    ansatz allowing higher-order poles.

  \item \textbf{General $n$ (Part~II, Theorem~\ref{thm:main}).}
    By strong induction, the same uniqueness result holds
    for all $n \geq 5$: cyclic $1$-zeros uniquely determine
    $\phi^3$ tree amplitudes. The induction uses the
    Homogeneity Lemma~\ref{lem:homogeneity},
    Dimensional Reduction (Proposition~\ref{prop:dimred}),
    and Inherited $1$-Zeros (Lemma~\ref{lem:inherited}).

  \item \textbf{Locality is derived, not assumed.}
    At each inductive step, finiteness of $B$ on the
    $1$-zero loci forces pole coefficients to zero,
    and the vanishing conditions kill the residual.
    Physical locality (poles only in planar channels)
    is a \emph{consequence} of the $1$-zero constraints
    together with the homogeneity of the ansatz space.

  \item \textbf{Independence of $1$-zero loci.}
    At $n = 5$, two or three loci suffice (depending on
    the ansatz). At $n = 6$, five of the six loci are
    needed. The earlier conjecture that $\lfloor n/2 \rfloor$
    loci suffice is falsified at $n = 6$; the correct
    count appears to be $n - 1$ (one locus is always
    redundant by cyclic symmetry).
\end{enumerate}

\subsection*{Remaining gap}

The proof of Theorem~\ref{thm:main} contains one flagged gap
in Step~4: the factorization argument for the BCFW residues
of a general $B \in \cB_n$ (not just $\cA_n^{\mathrm{tree}}$).
The claim that the residue at each BCFW pole factorizes as a
product of lower-point quantities, and hence vanishes when
$\widetilde{B}|_{\cZ_1} = 0$, relies on the combinatorial
correspondence between BCFW poles and sub-polygon factorisations
established in \cite[Sec.~3]{Rodina2024}. Transposing this
argument to the abstract ansatz setting $\cB_n$ requires
verifying that the pole structure of an arbitrary $B \in \cB_n$
admits the same sub-polygon decomposition as the Feynman-diagram
sum. At $n = 5$ and $n = 6$ this gap is bypassed by the
independent direct computation (Part~I and
Proposition~\ref{prop:n6}). For $n \geq 7$, either a direct
computation at $n = 7$ (feasible but computationally expensive)
or a self-contained algebraic proof of the factorization
structure would close this gap.

\subsection*{Open problems}

\begin{enumerate}[label=(\roman*)]
  \item Close the factorization gap in Step~4 by a
    self-contained algebraic argument, independent of
    \cite{Rodina2024}.

  \item Determine the sharp bound for the number of
    independent $1$-zero loci at general $n$.
    The data suggests $n - 1$; a proof would explain
    why exactly one locus is always redundant.

  \item Extend the framework to Yang--Mills and graviton
    amplitudes, where the ansatz space and $1$-zero
    structure are more intricate.
\end{enumerate}

\subsection*{Acknowledgements}

Computational verification was performed using
SymPy~\cite{SymPy2017} and NumPy; scripts are available in the
accompanying \texttt{computations/} directory.

\bigskip
\bibliography{references}
\bibliographystyle{unsrt}
